---
output: 
  github_document:
---

# surveyjoin

<!-- badges: start -->

[![R-CMD-check](https://github.com/DFO-NOAA-Pacific/surveyjoin/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DFO-NOAA-Pacific/surveyjoin/actions/workflows/R-CMD-check.yaml) [![DOI](https://zenodo.org/badge/484561620.svg)](https://zenodo.org/doi/10.5281/zenodo.10031852)

<!-- badges: end -->

This is a repository for combining trawl survey datasets from NOAA and Fisheries and Oceans Canada in the Northeast Pacific Ocean.

A pkgdown site is available [here](https://dfo-noaa-pacific.github.io/surveyjoin/).

This data includes surveys conducted by the Northwest Fisheries Science Center (NWFSC) off the west coast of the United States, surveys collected by Fisheries and Oceans Canada (DFO) in the waters of British Columbia, and surveys conducted by the Alaska Fisheries Science Center (AFSC) in Alaska.

```{r, include=FALSE, echo=FALSE}
sp <- surveyjoin::get_species()
```

There are `r nrow(sp)` species included in the initial version of the package, focusing on species that are occurring in multiple regions.

### Installing

```{r, eval=FALSE}
# install.packages("pak")
pak::pkg_install("DFO-NOAA-Pacific/surveyjoin")
```

### Basic use

```{r}
library(surveyjoin)
```

On first use, download the data and load it into a local SQL database:

```{r, eval=FALSE}
cache_data()
load_sql_data()
```

Find available species:

```{r}
get_species()
```

Load data for a species:

```{r}
d <- get_data("pacific cod")
```

```{r}
dplyr::glimpse(d, width = 72)
```

### Citations

Citing the `surveyjoin` package can be done with the DOI linked above, though more detailed citations may be needed for specific surveys or methodology. Some recent citations are:

[Aleutian Islands Bottom Trawl Survey]{.underline}

-   Von Szalay PG, Raring NW, Siple MC, Dowlin AN, Riggle BC, and Laman EA. 2023. Data Report: 2022 Aleutian Islands bottom trawl survey. U.S. Dep. Commer. DOI: 10.25923/85cy-g225.

[Gulf of Alaska Bottom Trawl Survey]{.underline}

-   Siple MC, von Szalay PG, Raring NW, Dowlin AN, Riggle BC. 2024. Data Report: 2023 Gulf of Alaska bottom trawl survey. DOI: 10.25923/GBB1-X748.

[Eastern & Northern Bering Sea Crab/Groundfish Bottom Trawl Surveys]{.underline}

-   Zacher LS, Richar JI, Fedewa EJ, Ryznar ER, Litzow MA. 2023. The 2023 Eastern Bering Sea Continental Shelf Trawl Survey: Results for Commercial Crab Species. U.S. Dep. Commer, 213 p.

-   Markowitz EH, Dawson EJ, Wassermann S, Anderson AB, Rohan SK, Charriere BK, Stevenson DE. 2024. Results of the 2023 eastern and northern Bering Sea continental shelf bottom trawl survey of groundfish and invertebrate fauna. U.S. Dep. Commer.

[Eastern Bering Sea Slope Bottom Trawl Survey]{.underline}

-   Hoff GR. 2016. Results of the 2016 eastern Bering Sea upper continental slope survey of groundfishes and invertebrate resources. U.S. Dep. Commer. DOI: 10.7289/V5/TM-AFSC-339.

[Fisheries and Oceans Canada Synoptic Bottom Trawl Surveys]{.underline}

-   Nottingham MK, Williams DC, Wyeth MR, Olsen N. 2017. *Summary of the West Coast Vancouver Island synoptic bottom trawl survey, May 28 – June 21, 2014*. DFO Can. Manuscr. Rep. Fish. Aquat. Sci. 2017/3140, viii + 55 p, Nanaimo.

-   Sinclair A, Schnute J, Haigh R, Starr P, Stanley R, Fargo J, Workman G. 2003. *Feasibility of Multispecies Groundfish Bottom Trawl Surveys on the BC Coast. DFO Canadian Science Advisory Secretariat (CSAS) Research Document, 2003/049.*

-   Williams DC, Nottingham MK, Olsen N, Wyeth MR. 2018a. *Summary of the Queen Charlotte Sound synoptic bottom trawl survey, July 6 – August 8, 2015*. DFO Can. Manuscr. Rep. Fish. Aquat. Sci. 3136, viii + 64 p, Nanaimo.

-   Williams DC, Nottingham MK, Olsen N, Wyeth MR. 2018b. *Summary of the West Coast Haida Gwaii synoptic bottom trawl survey, August 25 – October 2, 2014*. DFO Can. Manuscr. Rep. Fish. Aquat. Sci. 2018/3134, viii + 42 p, Nanaimo.

-   Wyeth MR, Olsen N, Nottingham MK, Williams DC. 2018. *Summary of the Hecate Strait synoptic bottom trawl survey, May 26 – June 22, 2015*. DFO Can. Manuscr. Rep. Fish. Aquat. Sci. 2018/3126, viii + 55 p, Nanaimo.

[USA West Coast Bottom Trawl Surveys]{.underline}

-   Keller AA, Wallace JR, Methot RD. 2017. The Northwest Fisheries Science Center’s West Coast Groundfish Bottom Trawl Survey: history, design, and description. DOI: 10.7289/V5/TM-NWFSC-136.

### What species are included?

We first divided the joined datasets into 4 major areas: Eastern Bering Sea, Gulf of Alaska, British Columbia, and west coast of California / Oregon / Washington states.

Within each region, we identified the species that occurred in at least 5% of all tows (resulting in 4 lists, 1 for each region). Because our interests are in cross-regional work, we identified species meeting our occurrence threshold that also occurred in 2 or more regions. This resulted in the following list of `r nrow(sp)` species:

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
sp |> 
  select(-itis) |> 
  arrange(common_name) |> 
  mutate(scientific_name = stringr::str_to_sentence(scientific_name)) |> 
  mutate(common_name = stringr::str_to_title(common_name)) |> 
  rename(`Scientific name` = scientific_name, `Common name` = common_name) |> 
  mutate(`Scientific name` = paste0("*", `Scientific name`, "*")) |> 
  knitr::kable()
```
